Sending a Balloon to the Edge of Space

### 2018-11-14 paper/paper.rmd
edits

There are four main layers to the atmosphere: the troposphere, stratosphere, mesosphere, and thermosphere. The **troposphere** is the section of the atmosphere from roughly 0km to 15km, and is the layer in which humans live. It includes 75% of the mass of all gases in the atmosphere, consisting of 78% nitrogen gas, 21% oxygen, and less than 1% of argon, carbon dioxide, and other constituents. Because it is both the lowest in altitude and most massive layer, the pressure in the troposphere is the greatest: 100kPa. As altitude increases in the troposphere, temperature decreases to about -60 \degree C. The** stratosphere** is the section from roughly 15km to 50km. The ozone layer, or the region in the atmosphere which absorbs most UV rays, is within the stratosphere at about 20km in altitude. The pressure in the stratosphere is 10kPa. The mesosphere is the region from 50km to 80km. The pressure in the mesosphere is 0.1 kPa. Finally, the thermosphere is the highest in altitude of the four layers, spanning from 80km into outer space. Here, the pressure is only 0.01kPa, as it is the least dense of the layers. The few air molecules in the thermosphere are heated to extremely high temperatures by the Sun’s radiation. However, because there are so few molecules, an appreciable difference in overall temperature cannot be measured, so the effective temperature of the thermosphere is less than that of the mesosphere.

@@

```{r cutdown_diagram, fig.cap="\\label{fig:figs} Cutdown mechanism circuit design"}
include_graphics("assets/cutdown.png")
```

@@

```{r simple_circuit}
include_graphics("assets/simple_circuit.png")
```
Figure 2: A simple example circuit

@@

The circuit consists of two Arduinos, one Mega ATK and one Uno. The Mega ATK is our primary microcontroller and controls the logger, transmitter, and receives input from every sensor except for the accelerometer. The accelerometer and another microSD logger are the only two components on the secondary Arduino Uno. The reason for the accelerometer needing it’s own Arduino is because the $I^2C$ library it required was not compatible with the other devices running on $I^2C$, such as the BME280 and spectrometer. The primary Arduino is powered by three batteries, and one battery powers the secondary arduino and the methane sensor. The methane sensor uses a large amount of power so it is with the secondary Arduino’s power supply so that if the sensor uses too much power it will not disrupt the majority of the data collection and sensors. All batteries go through 1 Amp 5 Volt voltage regulators, one for the primary Arduino and one for the secondary Arduino and methane sensor. The built in voltage regulator on the Arduino was not powerful enough and causing the sensors to malfunction so exterior regulators were included. Every sensor connects individually to either an analog pin or digital pin, or is one of the three sensors that connect to the SDA and SCL pins.

@@

```{r code_flow_chart, out.height="400px", fig.align='center'}
include_graphics("assets/flow_code.jpg")
```
Figure 3: Code Flow Chart

@@

grid.arrange(p1, p2, ncol=2)
```
Figure 5

@@

```{r vernier_cal2, fig.height=3.5, warning=F}
pressure3 <- read.csv("data/pressure3.csv")
names(pressure3)[1] <- "Vernier sensor (10-bit unit)"
names(pressure3)[2] <- "Extech sensor (kPa)"
p3 <- ggplot(pressure3, aes(x=pressure3$`Vernier sensor (10-bit unit)`, y=pressure3$`Extech sensor (kPa)`)) + geom_point() + labs(title="Pressure conversion between\n standard (Extech) and Arduino (Vernier) sensor", x="Pressure from Vernier (10-bit unit)", y="Pressure from Extech (kPa)") + geom_smooth(method = "lm") + geom_text(x=300, y=90, label=TeX("y = -25.3 + 0.255x, r^2 = 1"), parse=TRUE)
p3
```
Figure 6

@@

Below is example data used to calibrate the Veriner Pressure Sensor.

`r kable(pressure3, caption="Varying average pressure readings from Arduino (Vernier) vs. Standard (Extech) sensor")`

@@

humidity1 <- read.csv("data/humidity1.csv")
lm_eqn_h1 <- function(df){
    m <- lm(df$Arduino.... ~ df$Standard...., df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(coef(m)[1], digits = 3), 
              b = format(coef(m)[2], digits = 3), 
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));                 
}
h1 <- ggplot(humidity1, aes(x=humidity1$Arduino...., y=humidity1$Standard....)) + geom_point() + labs(title="Humidity conversion between\n standard (Vernier) and Arduino (BME280) sensor", x="Relative humidity from BME280 (% RH)", y="Relative humidity from Vernier (% RH)") + geom_smooth(method = "lm") + geom_text(x=55, y=60, label=TeX("y = 9.12 + 1.07x, r^2 = 0.997"), parse=TRUE)
h1
```
Figure 7

@@

```{r bme_temp, fig.height=4}
temp1 <- read.csv("data/temp1.csv")
lm_eqn_t1 <- function(df){
    m <- lm(df$Arduino...C. ~ df$Fieldpiece...C., df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(coef(m)[1], digits = 3), 
              b = format(coef(m)[2], digits = 3), 
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));                 
}
t1 <- ggplot(temp1, aes(x=temp1$Arduino...C., y=temp1$Fieldpiece...C.)) + geom_point() + labs(title="Temperature conversion between\n standard (Fieldpiece) and Arduino (BME280) sensor", x="Temperature from BME280 (°C)", y="Temperature from Fieldpiece's T1 rod (°C)") + geom_smooth(method = "lm") + geom_text(x=5, y=20, label=TeX("y = 0.122 + 1.03x, r^2 = 0.999"), parse=TRUE)
t1
errors <- read.csv("data/errors.csv")
names(errors)[2] <- "Percent error (%)"
```
Figure 8

@@

thermcal <- read.csv("data/temp2.csv")
temp <- thermcal$extech
cal <- lm(thermcal$thermistor ~ temp)
ggplot(thermcal, aes(x=temp, y=thermistor)) + geom_point() + geom_smooth(method="lm") + labs(title="Temperature Measured with Extech Sensor vs. Thermistor", x="Extech Measured Temperature (C)", y="Thermistor Measured Temperature")
```
Figure 9

@@

The data gathered from the pressure, humidity, and temperature sensors do not all accurately represent the universally true values. There is bound to be inaccuracy from at least one of the three sources of error: accuracy, resolution, and calibration of the sensor. The standard for pressure and temperature, the Extech SD700, has a very high accuracy of ±2 hPa and , which can be disregarded as insignificant. Using Graphical Analysis and the LINEST function in Excel, the uncertainty of the slope of Graph 7 was calculated. Thus, the percent error of the slope = uncertainty of slopeslope*100, which was later multiplied with all of the pressure readings collected by the sensor to calculate error bars. To calculate the error bars that either contain or depend on the humidity and temperature data, multiplying the percent error of the slope of the conversion graphs does not work because the accuracy of the standards could not be regarded as perfectly accurate. Instead of solely using the percent errors, the room mean square error (RMSE) was found for all sensors, as shown below in Table 5. Table 6 lists the total errors and percent errors for the four main experiments—one extra due to the addition of the thermistor calibration after the final launch.

**JOHN INSERT RMSE TABLE HERE**

>>>>

There are four main layers to the atmosphere: the troposphere, stratosphere, mesosphere, and thermosphere. The **troposphere** is the section of the atmosphere from roughly 0km to 15km, and is the layer in which humans live. It includes 75% of the mass of all gases in the atmosphere, consisting of 78% nitrogen gas, 21% oxygen, and less than 1% of argon, carbon dioxide, and other constituents. Because it is both the lowest in altitude and most massive layer, the pressure in the troposphere is the greatest: 100kPa. As altitude increases in the troposphere, temperature decreases to about -60 \degree C. The **stratosphere** is the section from roughly 15km to 50km. The ozone layer, or the region in the atmosphere which absorbs most UV rays, is within the stratosphere at about 20km in altitude. The pressure in the stratosphere is 10kPa. The mesosphere is the region from 50km to 80km. The pressure in the mesosphere is 0.1 kPa. Finally, the thermosphere is the highest in altitude of the four layers, spanning from 80km into outer space. Here, the pressure is only 0.01kPa, as it is the least dense of the layers. The few air molecules in the thermosphere are heated to extremely high temperatures by the Sun’s radiation. However, because there are so few molecules, an appreciable difference in overall temperature cannot be measured, so the effective temperature of the thermosphere is less than that of the mesosphere.

@@

```{r cutdown_diagram}
include_graphics("assets/cutdown.png")
```
Figure 2: Cutdown mechanism circuit design

@@

```{r simple_circuit}
include_graphics("assets/simple_circuit.png")
```
Figure 3: A simple example circuit

@@

The circuit consists of two Arduinos, one Mega ATK and one Uno. The Mega ATK is our primary microcontroller and controls the logger, transmitter, and receives input from every sensor except for the accelerometer. The accelerometer and another microSD logger are the only two components on the secondary Arduino Uno. The reason for the accelerometer needing it’s own Arduino is because the $I^2C$ library it required was not compatible with the other devices running on $I^2C$, such as the BME280 and spectrometer. The primary Arduino is powered by three batteries, and one battery powers the secondary arduino and the methane sensor. The methane sensor uses a large amount of power so it is with the secondary Arduino’s power supply so that if the sensor uses too much power it will not disrupt the majority of the data collection and sensors. All batteries go through 1 Amp 5 Volt voltage regulators, one for the primary Arduino and one for the secondary Arduino and methane sensor. The built in voltage regulator on the Arduino was not powerful enough and causing the sensors to malfunction so exterior regulators were included. Every sensor connects individually to either an analog pin or digital pin, or is one of the three sensors that connect to the SDA and SCL pins. See Appendix for schematic.

@@

```{r code_flow_chart, out.height="400px", fig.align='center'}
include_graphics("assets/flow_code.jpg")
```
Figure 4: Code Flow Chart

@@

grid.arrange(p1, p2, ncol=2)
```
Figure 5:

@@

```{r vernier_cal2, fig.height=3.5, warning=F}
pressure3 <- read.csv("data/pressure3.csv")
names(pressure3)[1] <- "Vernier sensor (10-bit unit)"
names(pressure3)[2] <- "Extech sensor (kPa)"
p3 <- ggplot(pressure3, aes(x=pressure3$`Vernier sensor (10-bit unit)`, y=pressure3$`Extech sensor (kPa)`)) + geom_point() + labs(title="Pressure conversion between\n standard (Extech) and Arduino (Vernier) sensor", x="Pressure from Vernier (10-bit unit)", y="Pressure from Extech (kPa)") + geom_smooth(method = "lm", color="purple") + geom_text(x=300, y=90, label=TeX("y = -25.3 + 0.255x, r^2 = 1"), parse=TRUE)
p3
```
Figure 6

@@

Below is example data used to calibrate the Veriner Pressure Sensor.
`r kable(pressure3, caption="Varying average pressure readings from Arduino (Vernier) vs. Standard (Extech) sensor")`

@@

humidity1 <- read.csv("data/humidity1.csv")
lm_eqn_h1 <- function(df){
    m <- lm(df$Arduino.... ~ df$Standard...., df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(coef(m)[1], digits = 3), 
              b = format(coef(m)[2], digits = 3), 
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));                 
}
h1 <- ggplot(humidity1, aes(x=humidity1$Arduino...., y=humidity1$Standard....)) + geom_point() + labs(title="Humidity conversion between\n standard (Vernier) and Arduino (BME280) sensor", x="Relative humidity from BME280 (% RH)", y="Relative humidity from Vernier (% RH)") + geom_smooth(method = "lm", color="purple") + geom_text(x=55, y=60, label=TeX("y = 9.12 + 1.07x, r^2 = 0.997"), parse=TRUE)
h1
```
Figure 7

@@

```{r bme_temp, fig.height=4}
temp1 <- read.csv("data/temp1.csv")
lm_eqn_t1 <- function(df){
    m <- lm(df$Arduino...C. ~ df$Fieldpiece...C., df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(coef(m)[1], digits = 3), 
              b = format(coef(m)[2], digits = 3), 
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));                 
}
t1 <- ggplot(temp1, aes(x=temp1$Arduino...C., y=temp1$Fieldpiece...C.)) + geom_point() + labs(title="Temperature conversion between\n standard (Fieldpiece) and Arduino (BME280) sensor", x="Temperature from BME280 (°C)", y="Temperature from Fieldpiece's T1 rod (°C)") + geom_smooth(method = "lm", color="purple") + geom_text(x=5, y=20, label=TeX("y = 0.122 + 1.03x, r^2 = 0.999"), parse=TRUE)
t1
errors <- read.csv("data/errors.csv")
names(errors)[2] <- "Percent error (%)"
```
Figure 8

@@

thermcal <- read.csv("data/temp2.csv")
temp <- thermcal$extech
cal <- lm(thermcal$thermistor ~ temp)
ggplot(thermcal, aes(x=temp, y=thermistor)) + geom_point() + geom_smooth(method="lm", color="purple") + labs(title="Temperature Measured with Extech Sensor vs. Thermistor", x="Extech Measured Temperature (C)", y="Thermistor Measured Temperature") + geom_text(x=-4, y=15, label=TeX("y = -12.89 + 1.499x, r^2 = 0.985"), parse=TRUE)
```
Figure 9

@@

The data gathered from the pressure, humidity, and temperature sensors do not all accurately represent the universally true values. There is bound to be inaccuracy from at least one of the three sources of error: accuracy, resolution, and calibration of the sensor. The standard for pressure and temperature, the Extech SD700, has a very high accuracy of ±2 hPa and , which can be disregarded as insignificant. Using Graphical Analysis and the LINEST function in Excel, the uncertainty of the slope of Graph 7 was calculated. Thus, the percent error of the slope = uncertainty of slope / slope * 100, which was later multiplied with all of the pressure readings collected by the sensor to calculate error bars. To calculate the error bars that either contain or depend on the humidity and temperature data, multiplying the percent error of the slope of the conversion graphs does not work because the accuracy of the standards could not be regarded as perfectly accurate. Instead of solely using the percent errors, the room mean square error (RMSE) was found for all sensors, as shown below in Table 5. Table 6 lists the total errors and percent errors for the four main experiments—one extra due to the addition of the thermistor calibration after the final launch.

### 2018-11-14 paper/paper.tex
edits

There are four main layers to the atmosphere: the troposphere,
stratosphere, mesosphere, and thermosphere. The \textbf{troposphere} is
the section of the atmosphere from roughly 0km to 15km, and is the layer
in which humans live. It includes 75\% of the mass of all gases in the
atmosphere, consisting of 78\% nitrogen gas, 21\% oxygen, and less than
1\% of argon, carbon dioxide, and other constituents. Because it is both
the lowest in altitude and most massive layer, the pressure in the
troposphere is the greatest: 100kPa. As altitude increases in the
troposphere, temperature decreases to about -60 \degree C. The**
stratosphere** is the section from roughly 15km to 50km. The ozone
layer, or the region in the atmosphere which absorbs most UV rays, is
within the stratosphere at about 20km in altitude. The pressure in the
stratosphere is 10kPa. The mesosphere is the region from 50km to 80km.
The pressure in the mesosphere is 0.1 kPa. Finally, the thermosphere is
the highest in altitude of the four layers, spanning from 80km into
outer space. Here, the pressure is only 0.01kPa, as it is the least
dense of the layers. The few air molecules in the thermosphere are
heated to extremely high temperatures by the Sun's radiation. However,
because there are so few molecules, an appreciable difference in overall
temperature cannot be measured, so the effective temperature of the
thermosphere is less than that of the mesosphere.

@@

\begin{figure}

{\centering \includegraphics[width=400px]{assets/cutdown}

}

\caption{\label{fig:figs} Cutdown mechanism circuit design}\label{fig:cutdown_diagram}
\end{figure}

@@

Figure 2: A simple example circuit

@@

@@

Figure 3: Code Flow Chart

@@

Figure 5

@@

The data gathered from the pressure, humidity, and temperature sensors
do not all accurately represent the universally true values. There is
bound to be inaccuracy from at least one of the three sources of error:
accuracy, resolution, and calibration of the sensor. The standard for
pressure and temperature, the Extech SD700, has a very high accuracy of
±2 hPa and , which can be disregarded as insignificant. Using Graphical
Analysis and the LINEST function in Excel, the uncertainty of the slope
of Graph 7 was calculated. Thus, the percent error of the slope =
uncertainty of slopeslope*100, which was later multiplied with all of
the pressure readings collected by the sensor to calculate error bars.
To calculate the error bars that either contain or depend on the
humidity and temperature data, multiplying the percent error of the
slope of the conversion graphs does not work because the accuracy of the
standards could not be regarded as perfectly accurate. Instead of solely
using the percent errors, the room mean square error (RMSE) was found
for all sensors, as shown below in Table 5. Table 6 lists the total
errors and percent errors for the four main experiments---one extra due
to the addition of the thermistor calibration after the final launch.

\textbf{JOHN INSERT RMSE TABLE HERE}

>>>>

There are four main layers to the atmosphere: the troposphere,
stratosphere, mesosphere, and thermosphere. The \textbf{troposphere} is
the section of the atmosphere from roughly 0km to 15km, and is the layer
in which humans live. It includes 75\% of the mass of all gases in the
atmosphere, consisting of 78\% nitrogen gas, 21\% oxygen, and less than
1\% of argon, carbon dioxide, and other constituents. Because it is both
the lowest in altitude and most massive layer, the pressure in the
troposphere is the greatest: 100kPa. As altitude increases in the
troposphere, temperature decreases to about -60 \degree C. The
\textbf{stratosphere} is the section from roughly 15km to 50km. The
ozone layer, or the region in the atmosphere which absorbs most UV rays,
is within the stratosphere at about 20km in altitude. The pressure in
the stratosphere is 10kPa. The mesosphere is the region from 50km to
80km. The pressure in the mesosphere is 0.1 kPa. Finally, the
thermosphere is the highest in altitude of the four layers, spanning
from 80km into outer space. Here, the pressure is only 0.01kPa, as it is
the least dense of the layers. The few air molecules in the thermosphere
are heated to extremely high temperatures by the Sun's radiation.
However, because there are so few molecules, an appreciable difference
in overall temperature cannot be measured, so the effective temperature
of the thermosphere is less than that of the mesosphere.

@@

\begin{center}\includegraphics[width=400px]{assets/cutdown} \end{center}

Figure 2: Cutdown mechanism circuit design

@@

Figure 3: A simple example circuit

@@

\begin{center}\includegraphics[width=400px]{assets/schematic} \end{center}

@@

Figure 4: Code Flow Chart

@@

Figure 5:

@@

The data gathered from the pressure, humidity, and temperature sensors
do not all accurately represent the universally true values. There is
bound to be inaccuracy from at least one of the three sources of error:
accuracy, resolution, and calibration of the sensor. The standard for
pressure and temperature, the Extech SD700, has a very high accuracy of
±2 hPa and , which can be disregarded as insignificant. Using Graphical
Analysis and the LINEST function in Excel, the uncertainty of the slope
of Graph 7 was calculated. Thus, the percent error of the slope =
uncertainty of slope / slope * 100, which was later multiplied with all
of the pressure readings collected by the sensor to calculate error
bars. To calculate the error bars that either contain or depend on the
humidity and temperature data, multiplying the percent error of the
slope of the conversion graphs does not work because the accuracy of the
standards could not be regarded as perfectly accurate. Instead of solely
using the percent errors, the room mean square error (RMSE) was found
for all sensors, as shown below in Table 5. Table 6 lists the total
errors and percent errors for the four main experiments---one extra due
to the addition of the thermistor calibration after the final launch.
